Estimating Neighborhood Rents with Scraped Data

### 2018-06-15 manuscript/index.Rmd
updating some language

This study demonstrates why scraped rental housing listings are a valuable source of data in subject areas such as housing studies and neighborhood demography. I draw on a set of descriptive analyses and models of quarterly rent levels among census tracts in Seattle, WA for this research objective. Summaries of the Craigslist data show agreement with existing sources in their distribution but do suggest higher median levels for most neighborhoods, likely given the listings' recent collection relative to other data. The models use a quarterly panel of observed neighborhood median one-bedroom rents and produce temporally- and spatially-smoothed predictions about neighborhood rent levels. The best model specification in terms of forecasting Q2 2018 based on Q1 2017 through Q1 2018 has a combination of local spatial smoothing and autoregression (order 1), with an average forecast error of about $140 or 8% relative to the median neighborhood. In a more general sense, the methods and frameworks used in this study also offer a template for efficiently collecting web data, conducting reproducible research, and publishing interactive results to GitHub.

I start by briefly reviewing some background relevant to studying neighborhood rental housing. I then outline a process for automated web scraping of a rental listing source like Craigslist using a program called Helena, and compare aggregates from the collected rental listings with other data to assess rent agreement in measures. After describing the data and process, I present a set of models predicting neighborhood rent levels to understand different how different temporal and spatial structures fit the data. I conclude by discussing the promise of web data for subfields like housing studies and the value of R Markdown and GitHub for social science research.

@@

I assume that the log of the median one-bedroom rent for Seattle neighborhoods is normally distributed and estimate the posterior distribution of parameters $\theta_{it}$ using [integrated nested Laplace approximations (INLA)](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf){target="_blank"} with the default, weakly-informative priors for all random-effects. I choose to log the dependent variable because all values are positive and this transformation improves model fit.

>>>>

This study demonstrates why scraped rental housing listings are a valuable source of data in subject areas such as housing studies and neighborhood demography. We draw on a set of descriptive analyses and models of quarterly rent levels among census tracts in Seattle, WA for this research objective. Summaries of the Craigslist data show agreement with existing sources in their distribution but do suggest higher median levels for most neighborhoods, likely given the listings' recent collection relative to other data. The models use a quarterly panel of observed neighborhood median one-bedroom rents and produce temporally- and spatially-smoothed predictions about neighborhood rent levels. The best model specification in terms of forecasting Q2 2018 based on Q1 2017 through Q1 2018 has a combination of local spatial smoothing and autoregression (order 1), with an average forecast error of about $140 or 8% relative to the median neighborhood. In a more general sense, the methods and frameworks used in this study also offer a template for efficiently collecting web data, conducting reproducible research, and publishing interactive results to GitHub.

We start by briefly reviewing some background relevant to studying neighborhood rental housing. I then outline a process for automated web scraping of a rental listing source like Craigslist using a program called Helena, and compare aggregates from the collected rental listings with other data to assess rent agreement in measures. After describing the data and process, We present a set of models predicting neighborhood rent levels to understand different how different temporal and spatial structures fit the data. We conclude by discussing the promise of web data for subfields like housing studies and the value of R Markdown and GitHub for social science research.

@@

We assume that the log of the median one-bedroom rent for Seattle neighborhoods is normally distributed and compute the posterior distributions of each model's hyperparameters and parameters $\theta_{it}$ using [integrated nested Laplace approximations (INLA)](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf){target="_blank"} with the default, weakly-informative priors for all random-effects. We choose to log the dependent variable because all values are positive and this transformation improves model fit.
